The state of knowledge for marine invertebrate biodiversity in the continental US

### 2017-08-28 manuscript.Rmd
better text about number of samples and diversity

<!-- add sentence -->
<!-- separate idigbio and (and add) obis in the results -->

We retrieved `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species (`r n_distinct(na.omit(combined_species$worms_phylum))` phyla) of marine invertebrates in the continental US EEZ by combining  `r format(nrow(idigbio_records), big.mark=",")` museum records from iDigBio (`r format(n_distinct(idigbio_records$worms_valid_name), big.mark = ",")` species, `r n_distinct(na.omit(idigbio_records$worms_phylum))` phyla), and `r format(nrow(obis_records), big.mark=",")` records from OBIS (`r format(n_distinct(obis_records$worms_valid_name), big.mark=",")` species,
`r n_distinct(na.omit(obis_records$worms_phylum))` phyla). Extant marine invertebrates occurring in the US represent less than 0.4% of the total iDigBio records, as it currently holds records for over 106 million specimens, and they represent about 1% of the records that occur in the United States (about 33 million records in iDigBio). Marine invertebrates that occur in the US represent approximately 4% of the 48 million occurrence records in OBIS.

<!-- need to add latitudinal gradient map -->
Most samples have been collected on the continental shelf (Fig. \@ref(fig:map-diversity-sampling)A). On the East coast, species richness is highest at low latitudes, while it increases towards high latitude on the West coast (Fig. \@ref(fig:map-diversity-sampling)B).

>>>>

<!-- separate idigbio and (and add) obis in the results -->

We retrieved `r format(n_distinct(combined_species$worms_valid_name), big.mark = ",")` species (`r n_distinct(na.omit(combined_species$worms_phylum))` phyla) of marine invertebrates in the continental US EEZ by combining  `r format(nrow(idigbio_records), big.mark=",")` museum records from iDigBio (`r format(n_distinct(idigbio_records$worms_valid_name), big.mark = ",")` species, `r n_distinct(na.omit(idigbio_records$worms_phylum))` phyla), and `r format(nrow(obis_records), big.mark=",")` records from OBIS (`r format(n_distinct(obis_records$worms_valid_name), big.mark=",")` species,
`r n_distinct(na.omit(obis_records$worms_phylum))` phyla). Extant marine invertebrates occurring in the US represent less than 0.4% of the total number of iDigBio records (106 million), and about 1% of the US records (33 million records). Marine invertebrates that occur in the US represent approximately 4% of the 48 million occurrence records in OBIS.

```{r local-richness}
count_spp <- . %>%
    dplyr::pull(worms_valid_name) %>%
    dplyr::n_distinct()

n_spp_gom <- combined_records %>%
    dplyr::filter(is_in_gom == TRUE) %>%
    count_spp

n_spp_east <- combined_records %>%
    dplyr::filter(is_east_coast == TRUE) %>%
    count_spp

n_spp_west <- combined_records %>%
    dplyr::filter(is_west_coast == TRUE) %>%
    count_spp
```

Most samples have been collected on the continental shelf (Fig. \@ref(fig:map-diversity-sampling)A). More records have been collected from the Atlantic coast, than the Gulf of Mexico, and the Pacific coast (`r format(sum(combined_records$is_east_coast), big.mark=",")`, `r format(sum(combined_records$is_in_gom), big.mark=",")`, and `r format(sum(combined_records$is_west_coast), big.mark=",")` records respectively). The highest species richness has been recorded from the Gulf of Mexico (`r format(n_spp_gom, big.mark=",")` species), followed by the Atlantic coast (`r format(n_spp_east, big.mark=",")` species) an d the Pacific coast (`r format(n_spp_west, big.mark=",")` species).

<!-- need to add latitudinal gradient map -->
On the East coast, species richness is highest at low latitudes, while it increases towards high latitude on the West coast (Fig. \@ref(fig:map-diversity-sampling)B).
